---
title: "Analysis of Kickstarter projects"
author: "Jesus Leal"
output: html_document
---

This project is an analysis of pledges for financing submitted to [kickstater](https://www.kickstarter.com/), a crowdfunding platform where 
entrepreneurs subtmit their ideas for funding. The dataset was obtained from 
[Kaggle](https://www.kaggle.com/kemical/kickstarter-projects). For this analysis
we will use libraries from the [Tidyverse](https://www.tidyverse.org/), a set of 
packages developed primarily by Hadley Wickham.

We start by loading the libraries.

``````{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(ggridges)
library(knitr)
library(lubridate)
```

Now that we have the workding directory we can just use the relative path of the 
GitHub repo where the data is also mantained. The next step in the analysis is
loading the data. The readr library is quite helpful here as the original file 
is in a zip format, but the library takes care of the heavylifting for us.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("/Users/jtrujillo/Dropbox/git")
knitr::opts_knit$set(root.dir = "/Users/jtrujillo/Dropbox/git")
kickstart_data<-
  read_csv('./udacity_dataanalyst/r_analysis_project/data/ks-projects-201801.csv.zip')
```

The data set was created by user Kermical via webscrape the plattform. It is 
composed of 15 variables and has 378K+ observations. The variables indicate the 
name of the project, the category of the project (e.g. food, restaurants, film 
and video, etc.), when was it launched, currency, number of backers, status of 
the project (succesful vs failed), etc.

The first variable I am interested in exploring is goal (funding requested).
This will provide a good idea on the range of projects that are pitched in the
plattform. We start this exploration with a histogram. To define the size of the
bins we use one half of the standard deviation of our variable of interest.

```{r message=FALSE, warning=FALSE}
summary(kickstart_data$goal)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
histogram_goal<- ggplot(kickstart_data, aes(x=goal))+
  geom_histogram(binwidth = 0.5*sd(kickstart_data$goal))+
  scale_x_continuous(labels=comma)+
  scale_y_continuous(labels=dollar_format(prefix="$"))+
  labs(x="Total funds requested per project (local currency)",
       y="Count of projects", title="Histogram of funds requested")+theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
histogram_goal

```

As we can see from the plot there a few observations with extremely large 
values. This can be explained by the fact that the goal variable is not 
standardized in a common currency, additionally there could be just a high
variation on the amount of money a given project requires. Let's explore how
the distribution looks if we used the standardized goal requested in USD.

```{r message=FALSE, warning=FALSE}
summary(kickstart_data$usd_goal_real)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
histogram_goal_usd<- ggplot(kickstart_data, aes(x=usd_goal_real))+
  geom_histogram(binwidth =sd(kickstart_data$usd_goal_real))+
  scale_x_continuous(labels=comma)+
  scale_y_continuous(labels=dollar_format(prefix="$"))+
  labs(x="Total funds requested per project (USD)",
       y="Count of projects", title="Funds Requested in USD")+
  theme(plot.title = element_text(hjust = 0.5))
histogram_goal_usd
```

We still have a large number of outliers in this case we will filter everything
that is above 1.5 times the interquartile range, defined as the difference 
between the 75th and 25th percentiles. If we follow this approach we end up with 
303,549 observations. This represents a loss of 19.7 percent of all the 
observations. To mitigate this problem I adopted a more lax definition of IQR 
and set the limit to 3.5 times the difference between the third and first 
quartile, with this criteria we have 341,563 observations, or 90 percent of the 
original dataset. The resulting histogram of funds requested without outliers
is below.

```{r echo=FALSE , message=FALSE, warning=FALSE}
observations_one_half<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_goal_real>1.5*(quantile(usd_goal_real,0.75)-
                                             quantile(usd_goal_real, 0.25)),
                        0,1))%>%filter(iqr_dummy==1)%>%
  summarise(total=sum(iqr_dummy))

observations_one_half<-observations_one_half[[1]]

observations_three_half<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_goal_real>3.5*(quantile(usd_goal_real,0.75)-
                                             quantile(usd_goal_real, 0.25)),
                        0,1))%>%filter(iqr_dummy==1)%>%
  summarise(total=sum(iqr_dummy))

observations_three_half<-observations_three_half[[1]]

#plot histogram filtering the outliers using the 3.5 IQR criteria
histogram_goal_usd_filtered<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_goal_real>3.5*(quantile(usd_goal_real,0.75)-
                                             quantile(usd_goal_real, 0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  ggplot(., aes(x=usd_goal_real))+
  geom_histogram(binwidth =1000)+
  scale_x_continuous(labels=dollar_format(prefix="$"))+
  scale_y_continuous(labels=comma)+
  labs(x="Total funds requested per project (USD)",
       y="Count of projects", title="Funds requested in USD (no outliers)")+ 
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
histogram_goal_usd_filtered
```

We now look at pledged funds (funds actully collected) in local currency. We 
also provide some summary statistics using the summary command. We will try to
assess if the funds pledge also show a high concentration on round numbers ( 
defined as those that fall on 5k, 10k, 20k, etc.)

```{r message=FALSE, warning=FALSE}
summary(kickstart_data$pledged)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
histogram_pledged<- ggplot(kickstart_data, aes(x=pledged))+
  geom_histogram(binwidth =sd(kickstart_data$pledged))+
  scale_x_continuous(labels=comma)+scale_y_continuous(labels=comma)+
  labs(x="Total funds obtained per project",
       y="Count of projects", title="Funds collected in local currency")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
histogram_pledged
```

In the previous plot we see that there are a some outliers that make the 
analysis hard withou any data transformation. We will see if this situation is 
also present once we look at fund raised in USD. As in the previous case we also
provide summary statistics with the summary commad. 

```{r message=FALSE, warning=FALSE}
summary(kickstart_data$usd_pledged_real)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
histogram_pledged_usd<- ggplot(kickstart_data, aes(x=usd_pledged_real))+
  geom_histogram(binwidth =2*sd(kickstart_data$usd_pledged_real))+
  scale_x_continuous(labels=dollar_format(prefix = "$"))+
  scale_y_continuous(labels=comma)+
  labs(x="Total funds requested per project (USD)",
       y="Count of projects", title="Funds collected in USD")+ theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
histogram_pledged_usd
```

We notice the same situation as with total the goal variable, a high presence of
outliers (the fact that we used USD did not make a significant difference). 
We follow the same approach and filter values higher than 3.5 timesthe IQR. This
leaves us with 341,526 observations. We now plot the histogram of projects by 
funding raised with no outliers in USD.

```{r echo=F, message=FALSE, warning=FALSE, echo=FALSE}
histogram_pledged_usd_filtered_three_half<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 
                                                        0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  summarise(total=sum(iqr_dummy))
histogram_pledged_usd_filtered_three_half<-
  histogram_pledged_usd_filtered_three_half[[1]]

histogram_pledged_usd_filtered<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 
                                                        0.25)),
                        1,0))%>%
  filter(iqr_dummy==0)%>%ggplot(., aes(x=usd_pledged_real))+
  geom_histogram(binwidth =1000)+
  scale_x_continuous(labels=dollar_format(prefix="$"))+
  scale_y_continuous(labels=comma)+
  labs(x="Total funds requested per project (USD)",
       y="Count of projects", title="Funds collected in UDS (no outliers)")+ 
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

histogram_pledged_usd_filtered
```

We can see that there are a lot of projects with no funds collected and no 
projects that collected money with round 'numbers' (5k, 10k, etc.). Let's 
explore the distribution of projects by their status (live, canceled, etc.).
In the next section of the analysis we will explore how status and projects
collected correlate. 

```{r echo=F, message=FALSE, warning=FALSE}

status_projects_bar_chart<-kickstart_data%>%group_by(state)%>%
  summarise(total=n())%>%
  arrange(-total)%>%
  ggplot(., aes(x=reorder(state,total), y=total))+geom_bar(stat='identity')+
  coord_flip()+theme_bw()+scale_y_continuous(labels=comma)+
  labs(x="Status of project",
       y="Count of projects", title="Projects by status")+
  theme(plot.title = element_text(hjust = 0.5))

status_projects_bar_chart

```

We can see there are a lot of projects  are failed (more than half). Let's now 
see how the project concentration changes by category (music, film, desing, 
etc.). This is an important step before we analyze correlations between 
different categorical variables.

```{r echo=F, message=F, warning=F}
category_projects_bar_chart<-kickstart_data%>%group_by(main_category)%>%
  summarise(total=n())%>%
  arrange(-total)%>%
  ggplot(., aes(x=reorder(main_category,total), y=total))+
  geom_bar(stat='identity')+
  coord_flip()+theme_bw()+scale_y_continuous(labels=comma, 
                                             limits = c(0,70000), 
                                             expand=c(0,0))+
  labs(x="Category of project",
       y="Count of projects", title="Projects by category")+
  theme(plot.title = element_text(hjust = 0.5))

category_projects_bar_chart
```

The most common category in the dataset corresponds to projects in the category 
film and video, followed closely by music, publishing and games. 

I am also interested in exploring how long are projects online. We did not get
this information on the original dataset but we will use the dates of when
the projects went online (launched) and when the dealine of the projects.

```{r message=F, echo=F, warning=F}
kickstart_data<-kickstart_data%>%
  mutate(days_online=difftime(deadline, launched, unit="days"))
kickstart_data$days_online<-as.numeric(kickstart_data$days_online)
histogram_days_project<-ggplot(kickstart_data, aes(x=days_online))+
  geom_histogram(binwidth = 5)+theme_bw()+labs(x="Days online",
       y="Count of projects", title="Projects by days online")+
  theme(plot.title = element_text(hjust = 0.5))

histogram_days_project
```

We can see there are some outliers due to parsing of the date. We filter those 
observations with a launch date with year equal to 1970. We repeat the same plot
this time without the project with wrong date formatting.

```{r, message=F, echo=F, warning=F}
histogram_days_project<-kickstart_data%>%
  filter(year(launched)>1970)%>%ggplot(., aes(x=days_online))+
  geom_histogram(binwidth = 5)+theme_bw()+labs(x="Total days a project was up",
       y="Count of projects", title="Days project online (no outlier)")+
  scale_x_continuous(breaks = seq(0, 100, by = 10), expand = c(0, 0), 
                     limits=c(0,95))+
  scale_y_continuous(labels=comma, expand = c(0, 0), limits = c(0,210000))+
  theme(plot.title = element_text(hjust = 0.5))

histogram_days_project  
```

We can see that the most common number of days for a project to be up is 30,
with 60 as the second most common option.

The last variable I am interested in exploring is the number of backers of 
projects and how they are distributed. In the bivariete section we will explore
how number of backers is correlated with success of a project. I also provide
summary statistics with the summary command.

```{r, message=F,  warning=F}
summary(kickstart_data$backers)
```

```{r, message=F, echo=F, warning=F}
backers_hist<-ggplot(kickstart_data, aes(x=backers))+
  geom_histogram(binwidth = 1000)+theme_bw()+labs(x="Total backers",
       y="Count of projects", title="Projects by backers")+
  theme(plot.title = element_text(hjust = 0.5))
backers_hist
```

As with the other variables we see that there are several outliers in the data.
We use the 3.5 IQR rule of thumb and plot the number of backers again.

```{r, echo=F, message=F, warning=F}
backers_hist_no<-kickstart_data%>%
mutate(iqr_dummy=ifelse(backers>3.5*(quantile(backers,0.75)-
                                               quantile(backers, 0.25)),
                        1,0))%>%
  filter(iqr_dummy==0)%>%ggplot(., aes(x=backers))+
  geom_histogram(binwidth =10)+
  scale_x_continuous(labels=comma, expand = c(0,0), breaks = seq(0, 220, 
                                                                 by = 10))+
  scale_y_continuous(labels=comma, expand = c(0,0), limits=c(0, 200000))+
  labs(x="Total backers per project",
       y="Count of projects", title="Projects by backers (no outliers)")+ 
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

backers_hist_no
```

The number of backers shows a distribution similar to funds obtained, with the
majority of projects having 0 backers.

Finally I will explore where are projects located. For this analysis I will use 
the variable country. We use a simple barchart.

```{r echo=F, message=F, warning=F}
country_projects_bar_chart<-kickstart_data%>%group_by(country)%>%
  summarise(total=n())%>%
  arrange(-total)%>%
  ggplot(., aes(x=reorder(country,total), y=total))+geom_bar(stat='identity')+
  coord_flip()+theme_bw()+scale_y_continuous(labels=comma, 
                                             limits = c(0,310000), 
                                             expand=c(0,0))+
  labs(x="Country of project",
       y="Count of projects", title="Projects by country")+
  theme(plot.title = element_text(hjust = 0.5))

country_projects_bar_chart
```

The most common place of origin of projects is the USA followed by GB and 
Canada.

#Analysis

### What is the structure of your dataset?

There are 378,661 variables on this dataset measuring different aspects of 
projects that were launched on the kickstart plattform. The most important
variables included in the dataset are:

* category: indicates the subcategory of the project. This is a character 
variable that needs to be transformed to a factor. This has more than 50 levels 
and can get very specific.

* main_category: indicates the main category of a project. This is a character 
variable that needs to be transformed to a factor. This has 14 levels. The most 
common category in the project is Film & Video, followed closely by Music.

* deadline and launched: These are the dates when the project were launched and
when the project was taken down. I created a new variable using the differences
between the two dates (days_online).

* goal: Amount of money the project wants to raise in local currency. It is a 
heavily skewed variable with some projects having a large value; the median 
project asked for $45,454 and the maximun amount asked was $16,6361,391. There 
is another variable that stardadizes all projects to USD (usd_goal_real). 

* pledged: Amount of money the project raised in local currency. It is a 
heavily skewed variable with some projects having a large value. There is 
another variable that stardadizes all projects to USD (usd_goal_real).

* country: This variable captures where the projects are located. This is a 
character variable that should be consider as factor. The vast majority of 
projects come from the US, with almost 300,000.

### What is/are the main feature(s) of interest in your dataset?

One of the most interesting parts of the dataset comes from the fact that the 
majority of projects (more than 50 percent) fail. Also another interesting 
aspect of the project is the fact that the majority of them are online for
30 days.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I am interested in exploring the influence of days a project is online, category
and subcategory in the amount of money and success of a project.

### Did you create any new variables from existing variables in the dataset?

I created a new variable to understand how long projects are live in the 
plattform (days_online). 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I had to omit outliers from the dataset to better visualize the distribution of
data. The alternative would have been to use a logarithmic transformation. This
approach was discarded since the distribution of the data did indeed follow a 
normal distribution. While we get information on the transformation and may use
this transformation on a regression model it is hard to appreciate the 
concentration of projects on a log scale.

##Bivariate analysis

I am also interested in exploring differences in money collected by status of
the project (succesful vs cancelled) and by category of project (film, food, 
etc.). So now I will explore the distribution of funding requested and funding
obtained first by status of project and then by category. We use the same 
criteria of filtering by 3.5 above the IQR. 

The first plot corresponds to funds requested by status. We use a series
of ridgeplots to better visualize this relationship.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10, fig.fullwidth=T}
funds_requested_status<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_goal_real>3.5*(quantile(usd_goal_real,0.75)-
                                               quantile(usd_goal_real, 0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  ggplot(.,aes(x=usd_goal_real, y = state, fill=state))+
  geom_density_ridges()+theme_bw()+
  scale_x_continuous(labels=dollar_format(prefix = "$"), expand = c(0.01,0))+
    scale_y_discrete(expand = c(0.01, 0))+
  scale_fill_discrete(guide = guide_legend(reverse=TRUE))+
  labs(x="Total funds requested by status of project",
       y="Distribution of projects", title="Funds requested by project status")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

funds_requested_status
  
```

The projects show a very similar distribution between status. We will explore 
now if funding collected by status shows any difference with funds requested. 
Again we use a ridgeplot.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10, fig.fullwidth=T}
funds_obtained_status<-kickstart_data%>%
mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 
                                                        0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  ggplot(.,aes(x=usd_pledged_real, y = state, fill=state))+
  geom_density_ridges()+theme_bw()+
  scale_x_continuous(labels=dollar_format(prefix = "$"), expand = c(0.01,0))+
    scale_y_discrete(expand = c(0.01, 0))+
  labs(x="Total funds obtained by status of project",
       y="Distribution of projects", 
       title="Funds collected by status of project")+ 
    theme_bw()+  scale_fill_discrete(guide = guide_legend(reverse=TRUE))+
      theme(plot.title = element_text(hjust = 0.5))
    

funds_obtained_status
```

The funds collected by status of project show litte difference with the 
exception of projects that were succesful.

Let's explote differences between funding requested and project category. I will 
use boxplots to have an easier time visualizing the distribution of funding by 
project category. We also filter outliers using the 3.5 IRQ rule mentioned 
previously.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10, fig.fullwidth=TRUE}
boxplot_goal_categories<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_goal_real>3.5*(quantile(usd_goal_real,0.75)-
                                               quantile(usd_goal_real, 0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  ggplot(., aes(x=main_category,y=usd_goal_real))+
                             geom_boxplot(aes(fill="#bab7b6"), 
                                          outlier.alpha = 0.1,
                                         outlier.colour = '#1967e5', 
                                         fatten=0.5)+
  geom_jitter(aes(colour="#1967e5"), 
    alpha=0.009, width = 0.5, height = 0.1)+
  scale_y_continuous(labels=dollar_format(prefix = "$"))+theme_bw()+
  labs(y="Requested funds by category project", x="Project Category",
       title="Funding requested by project category USD")+
  scale_color_manual(values=c('#1967e5'))+guides(fill=FALSE, colour=FALSE)+
  scale_fill_manual(values=c("#bab7b6"))+
  stat_summary(geom = "crossbar", width=0.65, fatten=1.5, color="white", 
               fun.data = function(x){c(y=median(x), ymin=median(x), 
                                        ymax=median(x))})+
  theme(plot.title = element_text(hjust = 0.5))
  
boxplot_goal_categories
```

We can see that there are some differences on the amount of money a project 
request by category. Projects on the design, food and technology have the
highst median value.

Now that we have seen the distribution by category and state of the project 
let's see how days_online, the variable created before, and funding pledged 
correlate.

```{r echo=FALSE, message=FALSE}
scatter_days_funding<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_goal_real>3.5*(quantile(usd_goal_real,0.75)-
                                               quantile(usd_goal_real, 0.25)),
                        1,0))%>%filter(iqr_dummy==0 & year(launched)>1970)%>%
  ggplot(., aes(x=days_online,y=usd_goal_real))+
  geom_point(alpha=0.01, color="#166cf7")+theme_bw()+
    scale_x_continuous(expand = c(0,0), 
                       breaks = seq(0, 150, by = 10))+
  scale_y_continuous(labels=dollar_format(prefix = "$"))+
  labs(y="Requested funds by category project", x="Project's day online",
       title="Requested funds by days online USD")+
  theme(plot.title = element_text(hjust = 0.5))

scatter_days_funding
```

We see that there is a high concentration of projects on the 30 days deadline 
and in the 60 days deadline. There seems to be no real correlation. 

Finally I want to explore the correlation between number of backers and money 
pledged by projects. As in previous plots we filter outliers.

```{r echo=FALSE, message=FALSE}
scatter_backers_funding<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 0.25)),
                        1,0),
         iqr_dummy_b=ifelse(backers>3.5*(quantile(backers,0.75)-
                                               quantile(backers, 0.25)),
                        1,0))%>%filter(iqr_dummy==0 & iqr_dummy_b==0)%>%
  ggplot(., aes(x=backers,y=usd_pledged_real))+
  geom_point(alpha=0.01, color="#166cf7")+theme_bw()+
    scale_y_continuous(labels=dollar_format(prefix = "$"), expand = c(0,0))+
  labs(y="Secured funds by category project", x="Project's backers", 
       title="Collected funds by number of backers USD")+
  theme(plot.title = element_text(hjust = 0.5))
  
scatter_backers_funding
```

# Bivariate Analysis
### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Before plotting some of the variables I was interested in understanding how 
different projects on the plattform differ on money raised and money they wanted
to raise. I was also interested in exploring how the money a project requested 
may relate to its status (live, cancelled, failed, etc.) I was also interested
in understanding if there was a clear relationship between number of backers of
a given project and how they relate to total funding secured.

Some of the relationships followed an expected pattern such as number of backers
and secured funds. The relationship there seems pretty linear, the more backers
more funds a project secured. 

Another relationship that showed an interesing pattern is the interplay between 
state of a project and actual funds secured. All the different categories have
a high concentration around zero dollars raised (cancelled projects, unknown, 
failed, etc.). The big exception is projects that actually did received funding
where we see a more variation on the amount of money raised.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

An interesting pattern I observed was the fact that projects in the design and
technology had the highest median value for funds requested. Given the nature
of film projects I expected this to be the highest category. 

Another interesting fact that I observed is the high concentration of funding 
requested around 'round' numbers such as 20k, 30k, 50k, etc. This type of 
relationship was also present in the the number of days a project was online and
the funds requested. 

### What was the strongest relationship you found?

The number of backers and the secured funding was a verys strong relationship. 
If we exclude the project with zero backers the relationship is quite clear.

# Multivariate Plots Section

Finally I wanted to explore the relationship between amount secured for a 
project, category of the project, and the state of the project.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10, fig.fullwidth=TRUE}
boxplot_pledged_categories<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 0.25)),
                        1,0))%>%filter(iqr_dummy==0)%>%
  ggplot(., aes(x=main_category,y=usd_pledged_real, color=state))+
                             #geom_boxplot(position=position_dodge(width=1.2))+
  geom_jitter(alpha=0.2)+
  scale_y_continuous(labels=comma)+theme_bw()+
  labs(y="Collected funds by category project", x="Project Category", 
       title="Funding by project category and status USD")+
  scale_color_manual(values=c("#0066ff","#ff0000", 
                              "#003300", "#348750", "#ed9a15", "#ff3399"))+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, hjust = 1))

boxplot_pledged_categories
```

We can see that there are a lot of failed projects for all categories (red 
dots). There are however some interesting patterns such as the relatively high
number of projects with more than 10k in funding for film and video and music.

I am also interested in exploring the relationship between secured funds 
(pledged), number of backers, and the state of the project.

```{r echo=FALSE, message=FALSE}
scatter_backers_funding_state<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,0.75)-
                                               quantile(usd_pledged_real, 0.25)),
                        1,0),
         iqr_dummy_b=ifelse(backers>3.5*(quantile(backers,0.75)-
                                               quantile(backers, 0.25)),
                        1,0))%>%filter(iqr_dummy==0 & iqr_dummy_b==0)%>%
  ggplot(., aes(x=backers,y=usd_pledged_real, color=state))+
  geom_point(alpha=0.2)+theme_bw()+
    scale_y_continuous(labels=dollar_format(prefix = "$"), expand = c(0,0))+
  scale_color_manual(values=c("#0066ff","#ff0000", "#f200ff", "#348750", 
                              "#ed9a15", "#0c090c"))+ 
  labs(y="Secured funds by category project", 
       x="Project's backers",
       title="Funds raised by by backers and status USD")+
    theme(plot.title = element_text(hjust = 0.5))

scatter_backers_funding_state
```

We can see that there is a clear correlation between backers, funding and state.
More backers a project has more funding and more likely it will be a succesful
project. One insteresting exception seems to be projects that were canceled that
nevertheless managed a decen amount of backers. Perhaps another combination of
variables such as days a project was online sheds light on this phenomenon.

In the next plot I explore days online, secured funding and project category.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=10, fig.fullwidth=T}
scatter_online_funding_cat<-kickstart_data%>%
  mutate(iqr_dummy=ifelse(usd_pledged_real>3.5*(quantile(usd_pledged_real,
                                                         0.75)-
                                               quantile(usd_pledged_real,
                                                        0.25)),
                        1,0),
         iqr_dummy_b=ifelse(days_online>3.5*(quantile(days_online,0.75)-
                                               quantile(days_online, 0.25)),
                        1,0))%>%filter(iqr_dummy==0 & iqr_dummy_b==0)%>%
  ggplot(., aes(x=days_online,y=usd_pledged_real, color=main_category))+geom_point(alpha=0.5)+theme_bw()+
    scale_y_continuous(labels=dollar_format(prefix = "$"), expand = c(0,0))+
  labs(y="Secured funds by category project", x="Project's backers", 
       title="Funds raised by number of days online and project category USD")+
    theme(plot.title = element_text(hjust = 0.5))

scatter_online_funding_cat
```

The plot does not seem to provide any helpful insights on a difference between 
days online and project category.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

It seems that if a project collects more than 30 backers its chances of success
are actually quite high; in the scatter plot of backers by funds collected by
state of the project this relationship is pretty clear. The majority of projects
that were succesful exhibit a linear relationship between backers and funds 
raised.

### Were there any interesting or surprising interactions between features?

More than interactions there were a few surprised on level of funding raised and
project category. Journalism projects had very few projects overall and very few
that were succesful. Design projects also showed an interesting pattern; a high 
number of failed projects even when they raised more than $5k. 

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

scatter_backers_funding_state

```

### Description One

The first plot that I wanted to explore sheds light on the relationship between
backers, state of the project and secured fundings. This plot was interesting 
since there seems to be a clear threshold of 50 backers and status of project;
after a project hits that number seems much more likely that the status will be
succesful. In this plot we can also see that the cancelled projects seem to have
issues not with the number of backers but with the actual funding collected. 
A large number of the cancelled projects seem to have 20+ backers but less than
$5,000 in funding. 

### Plot Two
```{r echo=FALSE, fig.width=10, fig.height=10, fig.fullwidth=TRUE, Plot_Two}
boxplot_pledged_categories
```

### Description Two

In this boxplot we can see the relationship between categories of projects, 
secured funding, and status of the project. It is interesting to see that for 
design there seems to be a high concentration of projects that have failed even
after getting funding levels above $5,000. To put this into perspective for 
theater there are not many projects with that level of funding that failed.
Another interesting relationship can be observed in the category Music
where there seems to be a lot of succesful projects with $10,000+ in funding.

### Plot Three
```{r echo=FALSE, Plot_Three}
funds_obtained_status
```

### Description Three

In this plot we can see that for all the categories of project the most common 
amount of resources raised is close to 0, even for live projects. We can also
see that projects that are succesful do not exhibit this level of concentration. 
------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

The kickstarter dataset provided some interesting insights on the dynamics of 
the crowdsourcing scene. One of the first and most interesting insights that I 
got from this dataset is the fact that the  majority of projects fail (51 
percent). There was another element that was quite perplexing and that is the 
fact that a lot of projects ask for round numbers for fuding (5k, 10k, 20k, etc.
). This is an interesting feature of the dataset that can also shed light on 
status of the project. 

The dataset had a lot of outliers for some of the key variables such as funding 
requested, funding raised, number of backers, and days online. To properly 
visualize and analize the data I had to filter observations that were above or 
below 3.5 times the IQR.

The vast majority of projects come from english speaking countries, with the USA
accounting for almost 300k out of the 378k projects in the dataset. Projects 
also exhibit a lot of variation by different categories. As mentioned  in the 
previous section, projects in the design and journalism space exhibit a very 
distinct pattern in terms of funding raised and status of project.

A limitation of the dataset is the lack of description of the projects, number 
of people involved as well as timeline to deliver the project. These variables 
would have allowed us additional datacuts that could shed light on what makes a
crowdsourced project succesful.